---
title: "Assignement 2"
author: "Sofia Davoli 813479"
date: "30/4/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(lpSolveAPI)
library(igraph)
```

# Problem 1

## 1 *Determine the nodes that will be visited by the BB algorithm and for each of them get the upper and lower limit deduced by the algorithm in the execution.*

B&B algorithm start from $P_{0}$, and split for $x_{1}$. 
if $x_{1}$=0 then $z_{best}$=9, the algorithm consider then the other split relative to $x_{2}$, $x_{3}$ but the solutions $z_{best}$=9 have no improvement or is not feasible. So in the left part B&B stop at $P_{2}$.

if $x_{1}$=1, then $z$=16.2 but the constraint are not satisfied, though other split are require for $x_{2}$ and $x_{3}$. In the split of $x_{2}$, 
if $x_{2}$=0, $z_{best}$ have no improvement,
if $z_{2}$=1, $z_{best}$ change to 16, the algorithm descend in this direction until the constraint of integrality are satisfied. 

B&B stop in  $P_{11}$ where $z_{best}$=14, x=(1,1,0,0) resulting in being the best solution for the problem.




Upper limit is obtain considering the integer part of the number in the previous split. (ex 17.8-->17)
Lower limit is obtain considering the possible minimun value the function can obtain in that precise split, with the previously assigned variables. (ex  considering the objective function $9x_{1}$ + $5x_{2}$ + $6x_{3} + $4x_{4}$ in split $P_{12}$ assigned variables are ($x_{1}$=1,$x_{2}$=0), other variables can still vary from 1-0. The Lowes limit is equal to 9x1 +5x0+6x0+4x0= 9)

limit for the split of the B&B algorithm are:

-$P_{1}$: upper 16 lower 9

-$P_{2}$: upper 16 lower 0

-$P_{12}$: upper 16 lower 9

-$P_{14}$: upper 13 lower 9

-$P_{7}$:upper 16 lower 14

-$P_{9}$ upper 16  lower 14


## 2 *Solve the problem with an ILP solver and check the value of the objective function matches the one
found at point 1.*

solving ILP problem with 0 constraint and 4 variables.

```{r, results= 'hide'}
model = make.lp(0,4) # 0 constraints, 4 variables
lp.control(model, sense="max") # original maximization problem
```

```{r}
set.objfn(model,obj=c(9,5,6,4)) # definition of the objective function using profit coef

# First costraint
add.constraint(model,
               xt=c(6,3,5,2),
               type="<=",rhs=10,
               indices=c(1,2,3,4)) 


# Second costraint
add.constraint(model,
               xt=c(1,1),
               type="<=",rhs=1,
               indices=c(3,4))


# Third costraint
add.constraint(model,
               xt=c(-1,1),
               type="<=",rhs=0,
               indices=c(1,3))

# Fourth costraint
add.constraint(model,
               xt=c(-1,1),
               type="<=",rhs=0,
               indices=c(2,4))


set.bounds(model, lower= c(0,0,0,0))# non negativity
set.type(model, c(1:4), "binary")
```

Solution

```{r}
solve(model)
get.variables(model)
get.objective(model)
```
The solution sorrespond to the one founded before.



# Problem 2


## 1 *Draw a network flow model to represent this problem.*
```{r}
el <- matrix( c("P","1",6.50, "E","1",7.50, "P","2",7,
"E","2",8, "P", "3", 8.25, "E","3", 7.25, "S", "3", 6.75, "E","4", 7.75, "S", "4", 7, "E","5", 7.50, "S", "5", 6.75),
nc = 3, byrow = TRUE)
g <- graph_from_edgelist(el = el[,1:2], directed = T)
edge_attr(g, "weight") <- el[,3]
plot(g, layout=layout_with_graphopt)
```

The length of the vector is proportional to it's cost.


## 2 *Implement your model and solve it.*

```{r, results='hide'}
edges= data.frame(index_i=c("P", "E", "P","E", "P","E", "S", "E", "S", "E", "S"),
 index_j =c(1,1,2,2,3,3,3,4,4,5,5),
 lb =c(0,0,0,0,0,0,0,0,0,0,0),
 ub =c(Inf,Inf,Inf,Inf,Inf,Inf,Inf,Inf,Inf,Inf,Inf),
 cost =c(6.50, 7.5, 7, 8, 8.25, 7.25, 6.75, 7.75, 7, 7.50, 6.75)) #variables DF
b = c(30000, 40000, 25000, 35000, 33000) #customer per region
model = make.lp(0,11) #possibles linkes 
lp.control(model, sense="min") #minimization problem
set.objfn(model,obj=edges$cost)

#constraint

set.bounds(model,lower=edges$lb,upper=edges$ub)

#link constraints
add.constraint(model,
 xt=c(1,1),
 type=">=",rhs=b[1],
 indices=c(1,2))

add.constraint(model,
 xt=c(1,1),
 type=">=",rhs=b[2],
 indices=c(3,4))

add.constraint(model,
 xt=c(1,1,1),
 type=">=",rhs=b[3],
 indices=c(5,6,7))

add.constraint(model,
 xt=c(1,1),
 type=">=",rhs=b[4],
 indices=c(8,9))

add.constraint(model,
 xt=c(1,1),
 type=">=",rhs=b[5],
 indices=c(10,11))

#capacity constraints
add.constraint(model,
 xt=c(1,1,1),
 type="<=",rhs=60000,
 indices=c(1,3,5))

add.constraint(model,
 xt=c(1,1,1,1,1),
 type="<=",rhs=70000,
 indices=c(2,4,6,8,10))

add.constraint(model,
 xt=c(1,1,1),
 type="<=",rhs=40000,
 indices=c(7,9,11))


set.type(model, c(1:11),type="integer") #model type
```

```{r}
solve(model)
```
```{r}
get.variables(model)
```
```{r}
get.objective(model)
```

## 3 *What is the optimal solution?*

```{r}
edges <- cbind(edges,y_val = get.variables(model))
edges[c("index_i", "index_j", "y_val")]
```

